\documentclass[main.tex]{subfiles}
\begin{document}
\section{Data}
\label{sec:data}
This section describes data used in the paper.


\subsection{Descriptives}
Estimation sample: 
All existing homeowners are native Danes (all household members are ethnically Danish). Interquartile range of the households age is less than 30 years in a neighborhood. Households who have at least stayed a year. 

Household where all members are in education, earn less than ~70,000 DKK or over 2,000,000 DKK. People that (likely) lives in kolleigiums. This is to ensure some degree homogeneity across homes. 

New Non-West (nearest) neighbors must be within 25m of the incumbent resident. 

\begin{enumerate}
    \item Map of "selected" \href{nabolagstlas.dk}{Nabolagsatlas} neighborhoods:
    \begin{itemize}
        \item Share of non-west inhabitants (1990 / 2010)
        \item Avg. no of non-west K-nearest neighbors (1990 / 2010) 
        \item Map of neighborhoods where the "experiment" occurs
    \end{itemize}
    \item Summary stats
    \begin{itemize}
        \item Dependent variable (move within 2 years of new K=3/4 nearest neighbor)
        \item Individual characteristics:
        \begin{itemize}
            \item Length of education
            \item Income (highest earner?)
            \item Age
            \item Employed
            \item ?
        \end{itemize}
        \item Neighborhood characteristics
        \begin{itemize}
            \item Median/average income,
            \item Population density
            \item Non-west share
            \item Native share
            \item No. of K=10-nearest non-west neighbors (?)
        \end{itemize}
        \item 3-4 columns (cond. on being from DK): All / new non-west neighbor (nearest) / new non-west neighbor (near)
    \end{itemize}
\end{enumerate}

\subsection{Balance test}

Before diving into the main results, I perform a series of balance tests to determine any potential (mean) differences between the "control" and "treatment" group. Consider the following equation:

\begin{equation}
    X_{i, j, t} = \phi_1 \mathbb{I}[r', k=n_{nearest}] + \phi_2 \mathbb{I}[r', k = n_{near}] + \phi_3 \mathbb{I}[r', k = n_{far}] + \gamma Z_{i, j, t} + \omega_{j, t} + \epsilon_{i, j, t} 
    \label{eq:balance_tests}
\end{equation}
Where $X$ is observables at the household/individual/neighborhood level(?). Eq. (\ref{eq:balance_tests}) is exactly the same as Eq. (\ref{eq:main_eq_schelling_behavior})
\end{document}
